---
title: "Data Import"
output: github_document
---

This file is for data imports.

# Load Packages
```{r}
library(tidyverse)
library(readxl)
library(haven)
library(ggplot2)
```

## Load first data set:
```{r}
## Import litter df:
litters_df =
  read_csv(file = "data/FAS_litters.csv", 
           na = c('.', "NA", ""))

## Change column labels to snake case.
litters_df = janitor::clean_names(litters_df)
```

## Load second data set:
```{r}
## Import pups df:
pups_df =
  read_csv("data/FAS_pups.csv", 
           skip = 3, 
           na = c('.', "NA", ""))

## Change column labels to snake case.
pups_df = janitor::clean_names(pups_df)
```

## `select` some variables:

```{r}
litters_df |> 
  select(group, litter_number)

litters_df |> 
  select(group:gd_of_birth)

litters_df |> 
  select(-pups_survive)

litters_df |> 
  select(starts_with("gd"), group) # reorders columns

litters_df |> 
  select(GROUP = group, everything())
```

## Rename and relocate data
```{r}
litters_df |> 
  rename(GROUP = group)

litters_df |> 
  relocate(litter_number, gd0_weight) # Moves columns to the front of the DF.
```


## In the pups data select litter #, sex, and PD ears.
```{r}
pups_df |> 
  select(litter_number, sex, pd_ears)
```

## `filter` some rows:
```{r}
litters_df |> 
  filter(gd_of_birth == 20, pups_born_alive != 6, pups_born_alive < 10)

litters_df |> 
  filter(group %in% c("Con7", "Con8"))

litters_df |> 
  drop_na()

litters_df |> 
  drop_na(gd0_weight)
```

# In the pups data, keep only rows where PD walk is < 11 and sex is 2:
```{r}
pups_df |> 
  filter(pd_walk < 11, sex == 2)
```

## `mutate` to change columns
```{r}
# Add and change some columns:
litters_df |> 
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    wt_gain_squared = wt_gain^2,
    group = str_to_lower(group)
  )
```

# In the pups dataset, create a var that subtracts 7 from PD walk:
```{r}
pups_df |> 
  mutate(
    pd_walk_minus_7 = pd_walk - 7
  )
```

# `arrange` data:
```{r}
litters_df |> 
  arrange(pups_born_alive)

litters_df |> 
  arrange(desc(pups_born_alive))
```

## Do multiple steps:
```{r}
litters_df = read_csv(file = "data/FAS_litters.csv", 
           na = c('.', "NA", "")) |> 
  janitor::clean_names() |> 
  select(group, starts_with("gd")) |> 
  drop_na() |> 
  mutate(
    wt_gain = gd18_weight - gd0_weight
  )
```

# Load pups, clean names, drop missing, keep group and pd vars and pd walk -7
```{r}
pups_df = read_csv(file = "data/FAS_pups.csv",
          skip = 3,
          na = c('.', "NA", "")) |> 
  janitor::clean_names() |> 
  select(litter_number, contains("pd")) |> 
  drop_na() |> 
  mutate(
    pd_walk_minus_7 = pd_walk - 7
  )
```

# Fitting a linear model:
```{r}
litters_df |> 
  lm(gd18_weight ~ gd0_weight, data = _) |> # add underscore to pipe data into the right place. 
  summary()
```

## Scatterplot
```{r}
litters_df |> 
  group_by(group) |> 
  summarize(gd0_weight_var = var(gd0_weight), gd18_weight_var = var(gd18_weight), mean_gain = mean(wt_gain)) |> 
  ggplot(aes(x = mean_gain, y = gd0_weight_var, color = group)) +
  geom_point() +
  geom_smooth(method = "lm")
```









